# Reproducible Research: Peer Assessment 1

 ```{r}
 library(plyr)
 library(lattice)
 ```{r}
  
 ## Loading and preprocessing the data
 ```{r}
 data <- read.csv("activity.csv", header=TRUE)
 cleaned <- data[which(data$steps != "NA"), ]
 ```
 ## What is mean total number of steps taken per day?
 ```{r}
 
 daily <- ddply(cleaned, .(date), summarise, steps=sum(steps))
 hist(daily$steps, main="Number of Steps", xlab="Total number of steps taken each day", col="green")
 
 mn <- mean(daily$steps)
 md <- median(daily$steps)
 ```
 The mean is `r mn` and the median is `r md`
 
 ## What is the average daily activity pattern?
 ```{r}
 avgdaily <- ddply(cleaned, .(interval), summarise, steps=mean(steps))
 plot(avgdaily$interval, avgdaily$steps, type="l", col="red", xlab="5-minute interval", ylab="Average number of steps taken",
      main="Average daily activity pattern")
 
 
 ```
 ## Entering missing values
 ```{r}
 
 # Total number of missing values in the dataset
 
 missing <- sum(is.na(data$steps))
 ```
 
 The total number of missing values is `r missing`
 ```{r}
 
 # Fill NA's with average for that 5-min interval
 
 avgdaily[avgdaily$steps==max(avgdaily$steps),]
 colnames(avgdaily)[2] <- "avgint"
 merged <- arrange(join(data, avgdaily), interval)
 
 # Create a new dataset that is equal to the original dataset but with the missing data filled in.
 
 merged$steps[is.na(merged$steps)] <- merged$avgint[is.na(merged$steps)]
 
 # Histogram
 
 dailynew <- ddply(merged, .(date), summarise, steps=sum(steps))
 hist(dailynew$steps, main="Number of Steps", xlab="Total number of steps taken each day", col="green",)
 
 # mean and median total number of steps taken per day
 
 mnnew <- mean(dailynew$steps)
 mdnew <- median(dailynew$steps) 
 ```
 The new mean is `r mnnew` and the median is `r mdnew`
 
 ## Are there differences in activity patterns between weekdays and weekends?
 ```{r}
 
 weekdays <- weekdays(as.Date(merged$date))
 wddata <- transform(merged, day=weekdays)
 wddata$wk <- ifelse(wddata$day %in% c("Saturday", "Sunday"),"weekend", "weekday")
 avgwk <- ddply(wddata, .(interval, wk), summarise, steps=mean(steps))
 
 xyplot(steps ~ interval | wk, data = avgwk, layout = c(1, 2), type="l" , col="red")
 ```
